A Structural and Functional Bioinformatics Study of QTY-designed Retinylidene Proteins

### 2025-05-11 Paper/Source/manuscript.tex
finalize first draft

\documentclass[fleqn, 10pt, lineno]{manuscript}
%%\usepackage{setspace}
%%\doublespacing
\usepackage{soul, hyperref}
\usepackage[utf8]{inputenc}

@@

\author[1]{Siqi Pan}
\author[2]{Shuguang Zhang}
\affil[1]{Shanghai World Foreign Language Academy, 400 Baihua Street, Shanghai 200233, China}
\affil[2]{Lab of Molecular Architecture, Media Lab, Massachusetts Institute of Technology, 77 Massachusetts Avenue, Cambridge, MA 02139, USA}

\corrauthor[2]{Shuguang Zhang}{Shuguang@MIT.EDU}

@@

Retinylidene proteins are a large family of retinal-binding light-sensitive proteins, found in many organisms ranging from microbes to human. Microbial opsins have been utilized in optogenetics, while animal opsins are essential for vision and other non-visual light-dependent metabolic functions. However, all retinylidene proteins are membrane proteins with hydrophobic transmembrane domains, which makes it challenging to study them. In this structural and functional bioinformatics study, we used the QTY (Glutamine, Threonine, Tyrosine) code to design water-soluble QTY analogs of retinylidene proteins, including 9 human opsins and 3 microbial opsins. We provide the superpositions of the AlphaFold3-predicted hydrophobic native proteins and their water soluble QTY variants, and experimentally-determined structures when available. We also provide a comparison of the surface hydrophobicity of the variants. Furthermore, we run molecular dynamics (MD) simulations of native and QTY-designed OPN2 and analyze their response to the 11-cis to all-trans isomerization of the retinal chromophore. The findings of this study indicate that the QTY code can be used as a robust tool to design water-solubility retinylidene proteins. These have potential applications in protein studies, therapeutic treatments, and bioengineering.

@@

Instead of taking a computational approach, we applied the QTY (Glutamine, Threonine, Tyrosine) code to systematically engineer water-soluble analogs with reduced hydrophobicity in membrane proteins. There are structural similarities between hydrophobic and polar amino acids: leucine (L) and glutamine (Q); isoleucine (I) / valine (V) and threonine (T); and phenylalanine (F) and tyrosine (Y), as can be observed on high-resolution electron density maps \citep{Zhang_2018, Zhang_2022, Tegler_2020}. This justifies the replacement of hydrophobic amino acids with polar ones. We initially applied the QTY code to design several detergent-free chemokines receptors, all of which retained structural thermal stability and native ligand-binding and enzymatic activities despite substantial changes to the transmembrane domain \citep{Zhang_2018, Tegler_2020}. We have also applied the QTY code to design water-soluble GPCRs, including chemokine receptors \citep{Zhang_2018, Qing_2019, Tegler_2020, Skuhersky_2021}, cytokine receptors \citep{Hao_2020}, and olfactory receptors \citep{Skuhersky_2021, Johnsson_2025}. In addition to GPCR, we have also used the QTY code to design water-soluble glucose transporters \citep{Smorodina_2022}, ABC transporters \citep{Pan_2024}, monoamine neurotransmitter transporters \citep{Karagol_Karagol_Alper_2024}, glutamate transporters \citep{Karagol_Karagol_Smorodina_2024, Karagol_2024}, mitochondrial megacomplex \citep{Chen_2025}, antibodies \citep{Li_Wang_2024}, potassium channels \citep{Smorodina_2024}, receptor kinases \citep{Li_Tang_2024}, transmembrane enzymes \citep{Chen_Pan_2025}, as well as bacterial enzymes with beta barrel structures \citep{Sajeev-Sheeja_2023, Sajeev-Sheeja_2024}. The QTY-designed water-soluble CXCR4 chemokine receptor has been successfully used in biomimetic sensors \citep{Qing_2023}. We have also applied the rQTY (reverse QTY) code to design self-assembled nanoparticles \citep{Meng_2023}.

@@

Regarding native OPN2, the membrane protein system was constructed using the web-based membrane builder CHARMM-GUI and was downloaded in GROMACS format \citep{Jo_2008, Wu_2014, Lee_2016}. The protein was centered in a rectangular box. The generated membrane models consisted of 40\% POPC, 40\% POPE, 10\% POPS, and 10\% cholesterol, which simulate a rod photoreceptor disk membrane \citep{Albert_2005}. The system was solvated in TIP3P water with 150 mM NaCl. 11-cis-retinal was manually added to the protein files. LINCS constraints were used for constraints, and Verlet integrator was used. Electrostatics was handled with Particle Mesh Ewald (PME), with both Coulomb and van der Waals interaction cutoffs set at 1.2 nm. The energy of the system was minimized using the steepest descent until the maximum forces converged below 1000 kJ/mol/nm. The standard six-step CHARMM-GUI NP$\gamma$T equilibration protocol \citep{Jo_2008} was used, with two 125-ps NVT equilibration simulations, and one 125-ps and three 500-ps NP$\gamma$T equilibration simulations. Temperature and pressure were maintained at 303.15 K and 1.0 bar, respectively, using the V-rescale thermostat and C-rescale barostat with surface-tension coupling. Following equilibration, a 10-ns production MD simulation was run. The retinal molecule was then manually switched to the all-trans state by rotating C13 to C15 and bonded hydrogen and methyl groups by 180 degrees around the C11-C12 bond. Energy minimization and equilibration were repeated exactly as before, and then a 120-ns production MD simulation was run.

@@

Although the chromophore, retinal, is bound to retinylidene proteins in an internal pocket, it is only incorporated into the protein after protein folding is complete. In other words, protein folding does not depend on the chromophore. Thus, it is justified to predict the ligand-free forms of the proteins. We used AlphaFold3 to predict the structure of the 9 native human opsins and their water-soluble QTY analogs. The structures superpose very well (Figure~\ref{fig:humansup} \textbf{a)} to \textbf{i)}). For OPN2, which has an available X-ray diffraction structure, we also superposed the experimentally determined structure with the AlphaFold3-predicted structures. Overall, the root mean square distances (RMSD) are small, from 0.307\AA to 0.611\AA, with only one exception of OPN2$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{EXP}}$ (RMSD = 0.999\AA). This shows that the water-soluble QTY analogs are quite similar to native proteins in terms of structure.

@@

We observe that QTY substitutions do not introduce changes to the essential NPxxY motif or the retinal-binding lysine of human retinylidene proteins. Therefore, we propose that QTY analogs of human opsins may have conserved functions and carry out a molecular dynamics (MD) simulation. We record and analyze the behavior of native OPN2 and its QTY analog before and after the 11-cis to all-trans isomerization.

@@

Zooming in on the retinal-binding pocket, we observe that retinal is held within its binding pocket by several forces: the counterions of the retinal Schiff base (RSB) (113E and 181E), the hydrophobic interactions with neighboring residues (for example, 265W, 268Y, and the 187--189 beta strand), and, sometimes, steric collisions (Figure~\ref{fig:binding}). Since retinal is a ligand that binds in an internal pocket, we were unable to calculate the Gibbs free energy of its binding. Nevertheless, the distribution of interaction energies are found to be roughly the same in both native OPN2 and the water-soluble QTY analog, which is a strong suggestion of functional conservation. Finally, we observe very different hydrogen bonding and water molecule patterns inside the binding pocket, which is mainly due to several F to Y substitutions in that region. This may cause retinal to interact with adjacent residues in a subtly different way, and may lead to a different absorption peak since the environment of the RSB, i.e., the binding pocket, determines the spectral characteristic of opsins \citep{Fenno_2011}.

@@

@@

Conceptualization: S.P., S.Z.; Formal analysis: S.P.; Investigation: Methodology: S.P.; Validation: S.P.; Data curation: S.P.; Writing—original draft preparation: S.P., S.Z.; Review and editing: S.P., S.Z.

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/sequences.jpg}
        \vspace{-0.5cm}
	\caption{\textbf{The protein sequence alignments of 12 retinylidene proteins and their QTY analogs. } Blue: human opsins; purple: microbial opsins. The symbols $|$ and $*$ indicate that amino acids are identical and different respectively. Amino acids L, V, I and F in TM (transmembrane) alpha helices are replaced with Q, T, Y and Y respectively. The alpha helices (blue) are shown above the protein sequences. The characteristics of the proteins are shown above the sequences. 
    The alignments are: 
    \textbf{a)} OPN1MW \textit{vs} OPN1MW$^{\textrm{QTY}}$, 
    \textbf{b)} OPN1LW \textit{vs} OPN1LW$^{\textrm{QTY}}$, 
    \textbf{c)} OPN1SW \textit{vs} OPN1SW$^{\textrm{QTY}}$, 
    \textbf{d)} OPN2 \textit{vs} OPN2$^{\textrm{QTY}}$, 
    \textbf{e)} OPN3 \textit{vs} OPN3$^{\textrm{QTY}}$, 
    \textbf{f)} OPN4 \textit{vs} OPN4$^{\textrm{QTY}}$, 
    \textbf{g)} OPN5 \textit{vs} OPN5$^{\textrm{QTY}}$, 
    \textbf{h)} RGR \textit{vs} RGR$^{\textrm{QTY}}$, 
    \textbf{i)} RRH \textit{vs} RRH$^{\textrm{QTY}}$, 
    \textbf{j)} BACR \textit{vs} BACR$^{\textrm{QTY}}$, 
    \textbf{k)} BACH \textit{vs} BACH$^{\textrm{QTY}}$, 
    \textbf{l)} ChR2 \textit{vs} ChR2$^{\textrm{QTY}}$. }
	\label{fig:sequences}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/superposition-human-pairwise.jpg}
	\caption{\textbf{Superposition of AlphaFold3-predicted native human retinylidene proteins, their QTY analogs, and experimentally determined structures. } For clarity, unstructured N- and C- terminal ends are deleted. 
    For \textbf{a)} to \textbf{i)}, green: AlphaFold3-predicted native structure; cyan: AlphaFold3-predicted QTY analog structure; magenta: experimentally determined structure. The superpositions are 
    \textbf{a)} OPN1MW$^{\textrm{AF3}}$ \textit{vs} OPN1MW$^{\textrm{QTY}}$, 
    \textbf{b)} OPN1LW$^{\textrm{AF3}}$ \textit{vs} OPN1LW$^{\textrm{QTY}}$, 
    \textbf{c)} OPN1SW$^{\textrm{AF3}}$ \textit{vs} OPN1SW$^{\textrm{QTY}}$, 
    \textbf{d)} OPN2$^{\textrm{AF3}}$ \textit{vs} OPN2$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{EXP}}$, 
    \textbf{e)} OPN3$^{\textrm{AF3}}$ \textit{vs} OPN3$^{\textrm{QTY}}$, 
    \textbf{f)} OPN4$^{\textrm{AF3}}$ \textit{vs} OPN4$^{\textrm{QTY}}$, 
    \textbf{g)} OPN5$^{\textrm{AF3}}$ \textit{vs} OPN5$^{\textrm{QTY}}$, 
    \textbf{h)} RGR$^{\textrm{AF3}}$ \textit{vs} RGR$^{\textrm{QTY}}$, 
    \textbf{i)} RRH$^{\textrm{AF3}}$ \textit{vs} RRH$^{\textrm{QTY}}$. 
    For \textbf{j)} and \textbf{k)}, green: OPN1MW; red: OPN1LW; blue: OPN1SW; purple: OPN2; cyan: OPN3; gray: OPN4; olive: OPN5; orange: RGR; pink: RRH. The superpositions are
    \textbf{j)} OPN1MW$^{\textrm{AF3}}$ \textit{vs} OPN1LW$^{\textrm{AF3}}$ \textit{vs} OPN1SW$^{\textrm{AF3}}$ \textit{vs} OPN2$^{\textrm{AF3}}$ \textit{vs} OPN3$^{\textrm{AF3}}$ \textit{vs} OPN4$^{\textrm{AF3}}$ \textit{vs} OPN5$^{\textrm{AF3}}$ \textit{vs} RGR$^{\textrm{AF3}}$ \textit{vs} RRH$^{\textrm{AF3}}$, 
    \textbf{k)} OPN1MW$^{\textrm{QTY}}$ \textit{vs} OPN1LW$^{\textrm{QTY}}$ \textit{vs} OPN1SW$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{QTY}}$ \textit{vs} OPN3$^{\textrm{QTY}}$ \textit{vs} OPN4$^{\textrm{QTY}}$ \textit{vs} OPN5$^{\textrm{QTY}}$ \textit{vs} RGR$^{\textrm{QTY}}$ \textit{vs} RRH$^{\textrm{QTY}}$. 
    }
	\label{fig:humansup}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/superposition-microbial.jpg}
	\caption{\textbf{Superposition of AlphaFold3-predicted native microbial retinylidene proteins, their QTY analogs, and experimentally determined structures. } For clarity, unstructured N- and C- terminal ends are deleted. 
    Green: AlphaFold3-predicted native structure; cyan: AlphaFold3-predicted QTY analog structure; magenta: experimentally determined structure. The superpositions are 
    \textbf{a)} BACR$^{\textrm{AF3}}$ \textit{vs} BACR$^{\textrm{QTY}}$ \textit{vs} BACR$^{\textrm{EXP}}$ monomer, 
    \textbf{b)} BACR$^{\textrm{AF3}}$ \textit{vs} BACR$^{\textrm{QTY}}$ \textit{vs} BACR$^{\textrm{EXP}}$ trimer, 
    \textbf{c)} BACH$^{\textrm{AF3}}$ \textit{vs} BACH$^{\textrm{QTY}}$ \textit{vs} BACH$^{\textrm{EXP}}$ monomer, 
    \textbf{d)} BACH$^{\textrm{AF3}}$ \textit{vs} BACH$^{\textrm{QTY}}$ \textit{vs} BACH$^{\textrm{EXP}}$ trimer, 
    \textbf{e)} ChR2$^{\textrm{AF3}}$ \textit{vs} ChR2$^{\textrm{QTY}}$ \textit{vs} ChR2$^{\textrm{EXP}}$ monomer, 
    \textbf{f)} ChR2$^{\textrm{AF3}}$ \textit{vs} ChR2$^{\textrm{QTY}}$ \textit{vs} ChR2$^{\textrm{EXP}}$ dimer. }
	\label{fig:microbialsup}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/hydrophobicity.jpg}
	\caption{\textbf{Hydrophobic surface of 12 retinylidene proteins and their water-soluble QTY analogs. } The native proteins have many hydrophobic (yellow) patches due to the presence of hydrophobic amino acids, including L, I, V, and F. After QTY substitution, hydrophilic Q, T, and Y have respectively replaced hydrophobic L, I/V, and F, and the hydrophobic patches in the surface of transmembrane helices have become more hydrophilic (cyan). For clarity, unstructured N- and C- terminal ends are deleted. 
    \textbf{a)} OPN1MW \textit{vs} OPN1MW$^{\textrm{QTY}}$, 
    \textbf{b)} OPN1LW \textit{vs} OPN1LW$^{\textrm{QTY}}$, 
    \textbf{c)} OPN1SW \textit{vs} OPN1SW$^{\textrm{QTY}}$, 
    \textbf{d)} OPN2 \textit{vs} OPN2$^{\textrm{QTY}}$, 
    \textbf{e)} OPN3 \textit{vs} OPN3$^{\textrm{QTY}}$, 
    \textbf{f)} OPN4 \textit{vs} OPN4$^{\textrm{QTY}}$, 
    \textbf{g)} OPN5 \textit{vs} OPN5$^{\textrm{QTY}}$, 
    \textbf{h)} RGR \textit{vs} RGR$^{\textrm{QTY}}$, 
    \textbf{i)} RRH \textit{vs} RRH$^{\textrm{QTY}}$, 
    \textbf{j)} BACR \textit{vs} BACR$^{\textrm{QTY}}$ monomer, 
    \textbf{k)} BACH \textit{vs} BACH$^{\textrm{QTY}}$ monomer, 
    \textbf{l)} ChR2 \textit{vs} ChR2$^{\textrm{QTY}}$ monomer, 
    \textbf{m)} BACR \textit{vs} BACR$^{\textrm{QTY}}$ trimer, 
    \textbf{n)} BACH \textit{vs} BACH$^{\textrm{QTY}}$ trimer, 
    \textbf{o)} ChR2 \textit{vs} ChR2$^{\textrm{QTY}}$ dimer. 
    }
	\label{fig:hydrophobicity}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/function-conformation.jpg}
	\caption{\textbf{The conformational changes of native OPN2 and its QTY analog before and after 11-cis to all-trans isomerization of the chromophore, retinal. } \textbf{a)} and \textbf{b)}: 1ns running averages of the root mean square distances (RMSD) of the protein-retinal complex, transmembrane helix 6 (TM6), the retinal-binding pocket, and retinal. By convention, the isomerization is set at time 0ns, which is indicated by a brown, vertical dashed line. \textbf{c)}: superpositions between cis-state OPN2, trans-state OPN2, cis-state OPN2$^{\textrm{QTY}}$, and trans-state OPN2$^{\textrm{QTY}}$. Blue: cis-state protein, orange: 11-cis-retinal; yellow: trans-state protein, greenish cyan: all-trans retinal. }
	\label{fig:conformation}
\end{figure}

>>>>

\documentclass[fleqn, 10pt]{manuscript}
%%\usepackage{setspace}
%%\doublespacing
\usepackage{soul, hyperref}
\usepackage[utf8]{inputenc}

@@

\author[1]{Siqi Pan}
\affil[1]{Shanghai World Foreign Language Academy, 400 Baihua Street, Shanghai 200233, China}

\corrauthor[1]{Siqi Pan}{siqipan2008@outlook.com}

@@

Retinylidene proteins are retinal-binding light-sensitive proteins found in organisms ranging from microbes to human. Microbial opsins have been utilized in optogenetics, while animal opsins are essential for vision and light-dependent metabolic functions. However, retinylidene proteins have hydrophobic transmembrane (TM) domains, which makes them challenging to study. In this structural and functional bioinformatics study, we used the QTY (Glutamine, Threonine, Tyrosine) code to design water-soluble QTY analogs of retinylidene proteins, including 9 human and 3 microbial opsins. We provide superpositions of the AlphaFold3-predicted hydrophobic native proteins and their water-soluble QTY analogs, and experimentally-determined structures when available. We also provide a comparison of surface hydrophobicity of the variants. Despite significant changes to the protein sequence (35.53\% to 50.24\% in the TM domain), protein characteristics and structures are well-preserved. Furthermore, we run molecular dynamics (MD) simulations of native and QTY-designed OPN2 (rhodopsin) and analyze their response to the isomerization of 11-cis-retinal to all-trans-retinal. The results show that the QTY analog has similar functional behavior to the native protein. The findings of this study indicate that the QTY code can be used as a robust tool to design water-solubility retinylidene proteins. These have potential applications in protein studies, therapeutic treatments, and bioengineering.

@@

Instead of taking a computational approach, we applied the QTY (Glutamine, Threonine, Tyrosine) code to systematically engineer water-soluble analogs with reduced hydrophobicity in membrane proteins. There are structural similarities between hydrophobic and polar amino acids: leucine (L) and glutamine (Q); isoleucine (I) / valine (V) and threonine (T); and phenylalanine (F) and tyrosine (Y), as can be observed on high-resolution electron density maps \citep{Zhang_2018, Zhang_2022, Tegler_2020}. This justifies the replacement of hydrophobic amino acids with polar ones. Zhang et al. initially applied the QTY code to design several detergent-free chemokines receptors, all of which retained structural thermal stability and native ligand-binding and enzymatic activities despite substantial changes to the transmembrane domain \citep{Zhang_2018, Tegler_2020}. They have also applied the QTY code to design water-soluble GPCRs, including chemokine receptors \citep{Zhang_2018, Qing_2019, Tegler_2020, Skuhersky_2021}, cytokine receptors \citep{Hao_2020}, and olfactory receptors \citep{Skuhersky_2021, Johnsson_2025}. In addition to GPCR, they have also used the QTY code to design water-soluble glucose transporters \citep{Smorodina_2022}, ABC transporters \citep{Pan_2024}, monoamine neurotransmitter transporters \citep{Karagol_Karagol_Alper_2024}, glutamate transporters \citep{Karagol_Karagol_Smorodina_2024, Karagol_2024}, mitochondrial megacomplex \citep{Chen_2025}, antibodies \citep{Li_Wang_2024}, potassium channels \citep{Smorodina_2024}, receptor kinases \citep{Li_Tang_2024}, transmembrane enzymes \citep{Chen_Pan_2025}, as well as bacterial enzymes with beta barrel structures \citep{Sajeev-Sheeja_2023, Sajeev-Sheeja_2024}. The QTY-designed water-soluble CXCR4 chemokine receptor has been successfully used in biomimetic sensors \citep{Qing_2023}. They have also applied the rQTY (reverse QTY) code to design self-assembled nanoparticles \citep{Meng_2023}.

@@

Regarding native OPN2, the membrane protein system was constructed using the web-based membrane builder CHARMM-GUI and was downloaded in GROMACS format \citep{Jo_2008, Wu_2014, Lee_2016}. The protein was centered in a rectangular box. The generated membrane models consisted of 40\% POPC, 40\% POPE, 10\% POPS, and 10\% cholesterol, which simulate a rod photoreceptor disk membrane \citep{Albert_2005}. The system was solvated in TIP3P water with 150 mM NaCl. 11-cis-retinal was manually added to the protein files. LINCS constraints were used for constraints, and Verlet integrator was used. Electrostatics was handled with Particle Mesh Ewald (PME), with both Coulomb and van der Waals interaction cutoffs set at 1.2 nm. The energy of the system was minimized using the steepest descent until the maximum forces converged below 1000 kJ/mol/nm. The standard six-step CHARMM-GUI NP$\gamma$T equilibration protocol \citep{Jo_2008} was used, with two 125-ps NVT equilibration simulations, and one 125-ps and three 500-ps NP$\gamma$T equilibration simulations. Temperature and pressure were maintained at 303.15 K and 1.0 bar, respectively, using the V-rescale thermostat and C-rescale barostat with surface-tension coupling. Following equilibration, a 10-ns production MD simulation was run. The retinal molecule was then manually switched to the all-trans state by rotating C13 to C15 and bonded hydrogen and methyl groups by 180 degrees around the C11-C12 bond. This mimics the effect of the absorption of a photon by 11-cis-retinal. Energy minimization and equilibration were repeated exactly as before, and then a 120-ns production MD simulation was run.

@@

Although the chromophore, retinal, is bound to retinylidene proteins in an internal pocket, it is only incorporated into the protein after protein folding is complete. In other words, protein folding does not depend on the chromophore. Thus, it is justified to predict the ligand-free forms of the proteins. We used AlphaFold3 to predict the structure of the 9 native human opsins and their water-soluble QTY analogs. The structures superpose very well (Figure~\ref{fig:humansup} \textbf{a)} to \textbf{i)}). For OPN2, which has an available X-ray diffraction structure, we also superposed the experimentally determined structure with the AlphaFold3-predicted structures. Overall, the root mean square distances (RMSD) are small, from 0.307{\AA} to 0.611\AA, with only one exception of OPN2$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{EXP}}$ (RMSD = 0.999\AA). This shows that the water-soluble QTY analogs are quite similar to native proteins in terms of structure.

@@

QTY substitutions do not introduce changes to the essential NPxxY motif or the retinal-binding lysine of human retinylidene proteins. Therefore, we propose that QTY analogs of human opsins may have conserved functions and carry out a molecular dynamics (MD) simulation. We record and analyze the behavior of native OPN2 and its QTY analog before and after the 11-cis to all-trans isomerization.

@@

Zooming in on the retinal-binding pocket, we observe that retinal is held within its binding pocket by several forces in both the native and QTY analog: the counterions of the retinal Schiff base (RSB) (113E and 181E), the hydrophobic interactions with neighboring residues (for example, 265W, 268Y, and the 187--189 beta strand), and, sometimes, steric collisions (Figure~\ref{fig:binding}). Since retinal is a ligand that binds in an internal pocket, we were unable to calculate the Gibbs free energy of its binding. Nevertheless, the distribution of interaction energies are found to be similar in native OPN2 and the water-soluble QTY analog, except for small differences: native OPN2 experiences a change in interaction energy at about $t=50$ns due to the entrance of water near 113E; the QTY analog has more negative interaction energy at residues 208 and 212 due to phenylalanine (F) to tyrosine (Y) substitutions. Besides from these, the similarity is a strong suggestion of functional conservation. Finally, we observe very different hydrogen bonding and water molecule patterns inside the binding pocket, which is mainly due to several F to Y substitutions in that region. This may cause retinal to interact with adjacent residues in a subtly different way, and may lead to a different absorption peak since the environment of the RSB, i.e., the binding pocket, determines the spectral characteristic of opsins \citep{Fenno_2011}.

@@

\section*{Supplementary Information}

The Supplementary Information is available at...

@@

Conceptualization: S.P.; Formal analysis: S.P.; Investigation: Methodology: S.P.; Validation: S.P.; Data curation: S.P.; Writing—original draft preparation: S.P.; Review and editing: S.P.

\section*{Acknowledgements}

S.P. thanks Prof. Shuguang Zhang of the Lab of Molecular Architecture, Media Lab, Massachusetts Institute of Technology for inspiring the topic of this study as well as providing valuable advice on research methods and manuscript preparation.

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/sequences.jpg}
	\caption{\textbf{The protein sequence alignments of 12 retinylidene proteins and their QTY analogs. } Blue: human opsins; purple: microbial opsins. The symbols $|$ and $*$ indicate that amino acids are identical or different, respectively. Amino acids L, I/V, and F in TM (transmembrane) alpha helices (shown in blue above the sequences) are replaced with Q, T, and Y respectively. The variation in the TM domain ranges from 35.53\% to 50.24\% while the overall variation rate ranges from 15.48\% to 30.04\%. The characteristics of the proteins are shown above the sequences. Despite large variation rates, the pI only experiences a slight change between 0.00 and 0.16 and the MW increases slightly by a value between 0.04kDa and 0.60kDa. The enlarged individual sequence alignments are available in Supplementary Information. 
    The alignments are 
    \textbf{a)} OPN1MW \textit{vs} OPN1MW$^{\textrm{QTY}}$, 
    \textbf{b)} OPN1LW \textit{vs} OPN1LW$^{\textrm{QTY}}$, 
    \textbf{c)} OPN1SW \textit{vs} OPN1SW$^{\textrm{QTY}}$, 
    \textbf{d)} OPN2 \textit{vs} OPN2$^{\textrm{QTY}}$, 
    \textbf{e)} OPN3 \textit{vs} OPN3$^{\textrm{QTY}}$, 
    \textbf{f)} OPN4 \textit{vs} OPN4$^{\textrm{QTY}}$, 
    \textbf{g)} OPN5 \textit{vs} OPN5$^{\textrm{QTY}}$, 
    \textbf{h)} RGR \textit{vs} RGR$^{\textrm{QTY}}$, 
    \textbf{i)} RRH \textit{vs} RRH$^{\textrm{QTY}}$, 
    \textbf{j)} BACR \textit{vs} BACR$^{\textrm{QTY}}$, 
    \textbf{k)} BACH \textit{vs} BACH$^{\textrm{QTY}}$, 
    \textbf{l)} ChR2 \textit{vs} ChR2$^{\textrm{QTY}}$. }
	\label{fig:sequences}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/superposition-human-pairwise.jpg}
	\caption{\textbf{Superposition of AlphaFold3-predicted native human retinylidene proteins, their QTY analogs, and experimentally determined structures. } For clarity, unstructured N- and C- terminal ends are deleted. 
    For \textbf{a)} to \textbf{i)}, despite significant changes in the protein sequence, the structures superpose very well. The root mean square distance (RMSD) values are quite small, from 0.307{\AA} to 0.611\AA, with only one exception (OPN2$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{EXP}}$, RMSD = 0.999\AA). Green: AlphaFold3-predicted native structure; cyan: AlphaFold3-predicted QTY analog structure; magenta: experimentally determined structure. The superpositions are 
    \textbf{a)} OPN1MW$^{\textrm{AF3}}$ \textit{vs} OPN1MW$^{\textrm{QTY}}$, 
    \textbf{b)} OPN1LW$^{\textrm{AF3}}$ \textit{vs} OPN1LW$^{\textrm{QTY}}$, 
    \textbf{c)} OPN1SW$^{\textrm{AF3}}$ \textit{vs} OPN1SW$^{\textrm{QTY}}$, 
    \textbf{d)} OPN2$^{\textrm{AF3}}$ \textit{vs} OPN2$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{EXP}}$, 
    \textbf{e)} OPN3$^{\textrm{AF3}}$ \textit{vs} OPN3$^{\textrm{QTY}}$, 
    \textbf{f)} OPN4$^{\textrm{AF3}}$ \textit{vs} OPN4$^{\textrm{QTY}}$, 
    \textbf{g)} OPN5$^{\textrm{AF3}}$ \textit{vs} OPN5$^{\textrm{QTY}}$, 
    \textbf{h)} RGR$^{\textrm{AF3}}$ \textit{vs} RGR$^{\textrm{QTY}}$, 
    \textbf{i)} RRH$^{\textrm{AF3}}$ \textit{vs} RRH$^{\textrm{QTY}}$. 
    For \textbf{j)} and \textbf{k)}, there is a large degree of similarity between the RMSD between a pair of native proteins and that between the corresponding pair of QTY analogs. Green: OPN1MW; red: OPN1LW; blue: OPN1SW; purple: OPN2; cyan: OPN3; gray: OPN4; olive: OPN5; orange: RGR; pink: RRH. The superpositions are
    \textbf{j)} OPN1MW$^{\textrm{AF3}}$ \textit{vs} OPN1LW$^{\textrm{AF3}}$ \textit{vs} OPN1SW$^{\textrm{AF3}}$ \textit{vs} OPN2$^{\textrm{AF3}}$ \textit{vs} OPN3$^{\textrm{AF3}}$ \textit{vs} OPN4$^{\textrm{AF3}}$ \textit{vs} OPN5$^{\textrm{AF3}}$ \textit{vs} RGR$^{\textrm{AF3}}$ \textit{vs} RRH$^{\textrm{AF3}}$, 
    \textbf{k)} OPN1MW$^{\textrm{QTY}}$ \textit{vs} OPN1LW$^{\textrm{QTY}}$ \textit{vs} OPN1SW$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{QTY}}$ \textit{vs} OPN3$^{\textrm{QTY}}$ \textit{vs} OPN4$^{\textrm{QTY}}$ \textit{vs} OPN5$^{\textrm{QTY}}$ \textit{vs} RGR$^{\textrm{QTY}}$ \textit{vs} RRH$^{\textrm{QTY}}$. 
    }
	\label{fig:humansup}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/superposition-microbial.jpg}
	\caption{\textbf{Superposition of AlphaFold3-predicted native microbial retinylidene proteins, their QTY analogs, and experimentally determined structures. } Despite significant changes in the protein sequence, the structures superpose very well. The root mean square distance (RMSD) values are quite small, with the highest being 0.685\AA. For clarity, unstructured N- and C- terminal ends are deleted. 
    Green: AlphaFold3-predicted native structure; cyan: AlphaFold3-predicted QTY analog structure; magenta: experimentally determined structure. 
    The superpositions are 
    \textbf{a)} BACR$^{\textrm{AF3}}$ \textit{vs} BACR$^{\textrm{QTY}}$ \textit{vs} BACR$^{\textrm{EXP}}$ monomer, 
    \textbf{b)} BACR$^{\textrm{AF3}}$ \textit{vs} BACR$^{\textrm{QTY}}$ \textit{vs} BACR$^{\textrm{EXP}}$ trimer, 
    \textbf{c)} BACH$^{\textrm{AF3}}$ \textit{vs} BACH$^{\textrm{QTY}}$ \textit{vs} BACH$^{\textrm{EXP}}$ monomer, 
    \textbf{d)} BACH$^{\textrm{AF3}}$ \textit{vs} BACH$^{\textrm{QTY}}$ \textit{vs} BACH$^{\textrm{EXP}}$ trimer, 
    \textbf{e)} ChR2$^{\textrm{AF3}}$ \textit{vs} ChR2$^{\textrm{QTY}}$ \textit{vs} ChR2$^{\textrm{EXP}}$ monomer, 
    \textbf{f)} ChR2$^{\textrm{AF3}}$ \textit{vs} ChR2$^{\textrm{QTY}}$ \textit{vs} ChR2$^{\textrm{EXP}}$ dimer. }
	\label{fig:microbialsup}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/hydrophobicity.jpg}
	\caption{\textbf{Hydrophobic surface of 12 retinylidene proteins and their water-soluble QTY analogs. } Hydrophobic patches are shown in yellow while hydrophilic patches are shown in cyan. The native proteins have many hydrophobic patches due to the presence of hydrophobic amino acids, including L, I, V, and F. After QTY substitution, hydrophilic Q, T, and Y have respectively replaced hydrophobic L, I/V, and F, and the hydrophobic patches in the surface of transmembrane helices have become more hydrophilic. In addition, the surface shape of the native and QTY analogs are very similar. For clarity, unstructured N- and C- terminal ends are deleted. 
    The comparisons are 
    \textbf{a)} OPN1MW \textit{vs} OPN1MW$^{\textrm{QTY}}$, 
    \textbf{b)} OPN1LW \textit{vs} OPN1LW$^{\textrm{QTY}}$, 
    \textbf{c)} OPN1SW \textit{vs} OPN1SW$^{\textrm{QTY}}$, 
    \textbf{d)} OPN2 \textit{vs} OPN2$^{\textrm{QTY}}$, 
    \textbf{e)} OPN3 \textit{vs} OPN3$^{\textrm{QTY}}$, 
    \textbf{f)} OPN4 \textit{vs} OPN4$^{\textrm{QTY}}$, 
    \textbf{g)} OPN5 \textit{vs} OPN5$^{\textrm{QTY}}$, 
    \textbf{h)} RGR \textit{vs} RGR$^{\textrm{QTY}}$, 
    \textbf{i)} RRH \textit{vs} RRH$^{\textrm{QTY}}$, 
    \textbf{j)} BACR \textit{vs} BACR$^{\textrm{QTY}}$ monomer, 
    \textbf{k)} BACH \textit{vs} BACH$^{\textrm{QTY}}$ monomer, 
    \textbf{l)} ChR2 \textit{vs} ChR2$^{\textrm{QTY}}$ monomer, 
    \textbf{m)} BACR \textit{vs} BACR$^{\textrm{QTY}}$ trimer, 
    \textbf{n)} BACH \textit{vs} BACH$^{\textrm{QTY}}$ trimer, 
    \textbf{o)} ChR2 \textit{vs} ChR2$^{\textrm{QTY}}$ dimer. 
    }
	\label{fig:hydrophobicity}
\end{figure}
